---
title: "STA445_Assignment3"
author: "Nicole Sylvester"
date: "2023-10-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(stringr)   # tidyverse string functions, not loaded with tidyverse 
library(refinr)    # fuzzy string matching
library(dplyr)
library(lubridate)
library(mosaicData)
```

## Chapter 11 Exercises  {#Exercises_Strings} 

1. For the following regular expression, explain in words what it matches on. Then add test strings to demonstrate that it in fact does match on the pattern you claim it does. Make sure that your test set of strings has several examples that match as well as several that do not. *If you copy the Rmarkdown code for these exercises directly from my source pages, make sure to remove the `eval=FALSE` from the R-chunk headers.*
    a) This regular expression matches: the pattern 'a' occurs in the strings
```{r}
strings <- c("hello", "goodbye", "apple")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'a') )
```
    b) This regular expression matches:  the pattern 'ab' occurs in the strings
```{r}
strings <- c("ab", "baba", "ba")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, 'ab') )
```
    c)  This regular expression matches: any string that has a or b in the string
```{r}
strings <- c("ab", "[ab]", "ba", "apple", "hello")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '[ab]') )
```
  d)  This regular expression matches:  strings that have a or b at the beginning of the string
```{r}
strings <- c("b", "ab", "a", "helloab" )
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '^[ab]') )
```
    e)  This regular expression matches:  strings that start with a digit, 
    followed by a white space, followed by an 'a' or 'A' then any other digits.
```{r}
strings <- c("1 a", "1a", "1 d", "2 Ab")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '\\d+\\s[aA]') )
```
    f)  This regular expression matches:  any string that starts with a digit,
    followed by zero or more spaces, followed by "a" or "A"
```{r}
strings <- c("5      ab", "5AB", "AB")
data.frame( string = strings ) %>%
mutate( result = str_detect(string, '\\d+\\s*[aA]') )
```

    g)  This regular expression matches:  strings that contain zero or 
    more digits and/or numbers.
```{r}
strings <- c("", " ", "3", "d ")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '.*') )
```
    h) This regular expression matches: Any String that starts with 
    2 repetitions of a alphanumeric charcter.
```{r}
strings <- c("aabar", "22bar", "22barabc", "23bar")
data.frame( string = strings ) %>%
    mutate( result = str_detect(string, '^\\w{2}bar') )
```

    i) This regular expression matches: any string that starts with the exact 
    string foo.bar OR a string that starts with two alphanumeric 
    characters followed by bar
```{r}
strings <- c("fobar", "aabar", "foo.bar", "foobar")
data.frame( string = strings ) %>%
  mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )
```
    

2. The following file names were used in a camera trap study. The S number represents the site, P is the plot within a site, C is the camera number within the plot, the first string of numbers is the YearMonthDay and the second string of numbers is the HourMinuteSecond.
    ```{r}
    file.names <- c( 'S123.P2.C10_20120621_213422.jpg',
                     'S10.P1.C1_20120622_050148.jpg',
                     'S187.P2.C2_20120702_023501.jpg')
    ```
    Produce a data frame with columns corresponding to the `site`, `plot`, `camera`, `year`, `month`, `day`, `hour`, `minute`, and `second` for these three file names. So we want to produce code that will create the data frame:
    
```{r, eval=FALSE}
     Site Plot Camera Year Month Day Hour Minute Second
     S123   P2    C10 2012    06  21   21     34     22
      S10   P1     C1 2012    06  22   05     01     48
     S187   P2     C2 2012    07  02   02     35     01
```

```{r}
df <- data.frame(
  file_info = file.names) %>%
  cbind(str_split_fixed(.$file_info, pattern='[\\.\\_]', n=6)) %>%
  rename(Site= `1`, Plot = `2`, Camera= `3`, Date = `4`, Time = `5` ) %>%
  mutate(
    Year = str_sub(Date, 1, 4),
    Month = str_sub(Date, 5, 6),
    Day = str_sub(Date, 7, 8)
  ) %>%
  select(-Date) %>% #remove date
   mutate(
   Hour = str_sub(Time, 1, 2),
   Minute = str_sub(Time, 3, 4),
   Second = str_sub(Time, 5, 6)
  ) %>%
  select(-Time) %>% #remove time
  select(-`6`) %>% #remove jpg
  select(-file_info) #remove file name

# Print the data frame
df
```



3. The full text from Lincoln's Gettysburg Address is given below. Calculate the mean word length *Note: consider 'battle-field' as one word with 11 letters*).
```{r}
Gettysburg <- 'Four score and seven years ago our fathers brought forth on this 
continent, a new nation, conceived in Liberty, and dedicated to the proposition 
that all men are created equal.

Now we are engaged in a great civil war, testing whether that nation, or any 
nation so conceived and so dedicated, can long endure. We are met on a great 
battle-field of that war. We have come to dedicate a portion of that field, as 
a final resting place for those who here gave their lives that that nation might 
live. It is altogether fitting and proper that we should do this.

But, in a larger sense, we can not dedicate -- we can not consecrate -- we can 
not hallow -- this ground. The brave men, living and dead, who struggled here, 
have consecrated it, far above our poor power to add or detract. The world will 
little note, nor long remember what we say here, but it can never forget what 
they did here. It is for us the living, rather, to be dedicated here to the 
unfinished work which they who fought here have thus far so nobly advanced. It 
is rather for us to be here dedicated to the great task remaining before us -- 
that from these honored dead we take increased devotion to that cause for which 
they gave the last full measure of devotion -- that we here highly resolve that 
these dead shall not have died in vain -- that this nation, under God, shall 
have a new birth of freedom -- and that government of the people, by the people, 
for the people, shall not perish from the earth.'
```

```{r}

text <- c(Gettysburg)

df<-data.frame(string = text) %>%
  mutate(result = str_split(string, "[\\,\\.\\s\\-{2}]+"))

wordMean <- str_length(df$result[[1]])
numStrings <- length(df$result[[1]])
wordAvg <- sum(wordMean)/numStrings

wordAvg
```


## Chapter 12 Exercises  {#Exercises_Dates}

1. Convert the following to date or date/time objects.
    a) September 13, 2010.
```{r}
mdy('September 13, 2010.')
```
    
    b) Sept 13, 2010.
```{r}
mdy('Sep 13, 2010.')
```
This does not work for Sept, only Sep


    c) Sep 13, 2010.
```{r}
mdy('Sep 13, 2010.')
```
    
    d) S 13, 2010. Comment on the month abbreviation needs.
```{r}
mdy('Sep 13, 2010') 
```
This does not work with only S. September must be abbreviated with 'Sep'
    
    e) 07-Dec-1941.
```{r}
dmy("07-Dec-1941.")

```
    
    f) 1-5-1998. Comment on why you might be wrong.
```{r}
dmy('1-5-1998.')
```
 This might be wrong because we don't know if 1 or 5 is the month or day.
 
 
    g) 21-5-1998. Comment on why you know you are correct.
```{r}
dmy('21-5-1998.')
```
This is correct because 21 cannot be a month number, so it must be a day. 
That means 5 must be the month.

    h) 2020-May-5 10:30 am
```{r}
ymd_hm('2020-May-5 10:30 am')
```
    
    i) 2020-May-5 10:30 am PDT (ex Seattle)
```{r}
ymd_hm('2020-May-5 10:30 am', tz='US/Pacific')
```
    
    j) 2020-May-5 10:30 am AST (ex Puerto Rico)
```{r}
ymd_hm('2020-May-5 10:30 am', tz='America/Puerto_Rico')
```
    
    

2. Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following _Write your code in a manner that the code will work on any date after you were born._:
    a) Calculate the date of your 64th birthday.
```{r}
dob <- mdy("May 15, 2002")

birthday64 <- dob + years(64)
birthday64
```

    b) Calculate your current age (in years). _Hint: Check your age is calculated correctly if your birthday was yesterday and if it were tomorrow!_ 
```{r}
dob <- mdy("May 15, 2002")
today <- Sys.Date()

difference <-interval(dob, today)

currentAge <-as.numeric(as.duration(difference), 'years')
currentAge
```

    d) Using your result in part (b), calculate the date of your next birthday.
```{r}
nextYear <- currentAge + 1
birthDate <- as.Date("2002-05-15")

nextAge <- as.numeric(format(birthDate, "%Y")) + nextYear

nextBirthday <- make_date(year=nextAge, month = 5, day = 15)

nextBirthday
```

    e) The number of _days_ until your next birthday.
```{r}
days <- as.numeric(nextBirthday - Sys.Date())
days
```
    
    f) The number of _months_ and _days_ until your next birthday.
```{r}
months <- floor(as.numeric(days)/30)
days <- days - months *30
cat("Months:", months, "Days:", days)

```
    

3. Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 
```{r}
AZTime <- ymd_hms("2015-05-08 15:00:00", tz = "America/Phoenix")

aucklandTime <- with_tz(AZTime, tz = "Pacific/Auckland")

aucklandTime

```


5. It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

    a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 
```{r}
data("Births78")
birthsData <- Births78[, c("date", "births")]
head(birthsData)
```

    b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?
```{r}
ggplot(data = birthsData, aes(x = date, y = births)) +
  geom_point() +
  labs(x = "Date", y = "Number of Births")
```

The number of births rise and fall based on dates most likely because of holidays, and seasonal trends.
    
    c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column. 
    
```{r}
birthsData <- birthsData %>%
  mutate(dow = wday(date, label = TRUE, abbr = FALSE))

# View the updated data frame
head(birthsData)
```

    d. Plot the data with the point color being determined by the day of the week variable.
```{r}
ggplot(data = birthsData, aes(x = date, y = births, color = dow)) +
  geom_point() +
  labs(x = "Date", y = "Number of Births", color = "Day of the Week")
```
    
    